Multifunctional Lipid Probes Background

### 2025-01-24 LipidProbe/PhosphatidicAcid.qmd
fleshing out overview

#### Study details

##### Authors
Alix Thomas, Rainer Mueller, Ana Kojic, Scotland E. Farley, Frank Stein, Per Haberkant, Carsten Schultz

##### Journal
Chemical Communications

@@

##### Abstract
Phosphatidylethanolamine (PE) is a ubiquitous lipid species in higher eukaryotes which resides preferentially in mitochondria. Here, we synthesized a multifunctionalized PE derivative (1) de-signed to identify PE-binding proteins in intact cells through photo-crosslinking and subsequent isolation and proteomic analysis of the PE-protein conjugates. Due to its aromatic caging group, 1 is initially concentrating in perinuclear membranes. After uncaging, rapid translocation to mitochondria was observed. Hence, the tool is useful for tracking PE location and for determining the PE interactome. A trifunctional phosphatidic acid (PA) derivative in compari-son was rapidly metabolized and is hence more limited in its use.

##### Lipid probes utilized

@@

##### Cell line analyzed

@@

##### Mass spectrometry quantification method

@@

##### Additional sample preparation <span class="help-symbol" title="e.g. sub-cellular fractionation, depletion, etc.">?</span>
:::

>>>>

## Study details

#### Authors
Alix Thomas, Rainer Mueller, Ana Kojic, Scotland E. Farley, Frank Stein, Per Haberkant, Carsten Schultz

#### Journal
Chemical Communications

@@

#### Abstract
Phosphatidylethanolamine (PE) is a ubiquitous lipid species in higher eukaryotes which resides preferentially in mitochondria. Here, we synthesized a multifunctionalized PE derivative (1) de-signed to identify PE-binding proteins in intact cells through photo-crosslinking and subsequent isolation and proteomic analysis of the PE-protein conjugates. Due to its aromatic caging group, 1 is initially concentrating in perinuclear membranes. After uncaging, rapid translocation to mitochondria was observed. Hence, the tool is useful for tracking PE location and for determining the PE interactome. A trifunctional phosphatidic acid (PA) derivative in compari-son was rapidly metabolized and is hence more limited in its use.

#### Lipid probes utilized

@@

#### Cell line analyzed

@@

#### Mass spectrometry quantification method

@@

::: {.callout-note collapse="true"}
##### ProteomeXchange details

**ProteomeXchange title:** Trifunctional lipid derivatives: PE’s mi-tochondrial interactome

**ProteomeXchange accession:** PXD059960

**PubMed ID:** Not applicable

**Publication DOI:** 10.1039/D4CC03599B

**Project Webpage:** http://www.ebi.ac.uk/pride/archive/projects/PXD059960

**FTP Download:** https://ftp.pride.ebi.ac.uk/pride/data/archive/2025/01/PXD059960
:::

#### Additional sample preparation <span class="help-symbol" title="e.g. sub-cellular fractionation, depletion, etc.">?</span>
:::

### 2025-01-24 TrifunctionalLipidProbesOverview.qmd
fleshing out overview

---
title: "Multifunctional Lipid Probes Background"
author: "Gaelen Guzman"
format: 
  html:
    embed-resources: false
standalone: true
---

@@

@@

For these reasons, synthetic analogs to native lipids have been a great utility toward probing the biological function of select lipid species. The chief modifications described in this repository

:::{.border}
```{r setup, echo=FALSE, include=FALSE}
library(plotly)

library(plotly)

@@

# Define hover regions (as rectangles)
shapes <- list(
  list(type = "rect", x0 = 0, x1 = 32, y0 = 50, y1 = 85,
       line = list(color = "red", width = 2),
       fillcolor = "rgba(255, 0, 0, 0.1)",  # Transparent fill
       name = "Region 1", hoverinfo = "none"),
  list(type = "rect", x0 = 50, x1 = 70, y0 = 40, y1 = 60,
       line = list(color = "blue", width = 2),
       fillcolor = "rgba(0, 0, 255, 0.1)",  # Transparent fill
       name = "Region 2", hoverinfo = "none")
)

# Define hover annotations
hover_annotations <- data.frame(
  x = c(20, 60),  # Midpoints of rectangles
  y = c(80, 50),
  text = c("Region 1 Info", "Region 2 Info")
)

# Create the plot
plot <- plot_ly() %>%
  add_trace(
    data = hover_annotations,
    x = ~x, y = ~y, text = ~text,
    mode = "markers",
    marker = list(opacity = 0),  # Invisible markers
    hoverinfo = "text"  # Show hover text
  ) %>%
  layout(
    shapes = shapes,
    images = list(
      list(
        source = image_url,
        xref = "x",
        yref = "y",
        x = 0,
        y = 100,
        sizex = 100,
        sizey = 100,
        xanchor = "left",
        yanchor = "top",
        layer = "below"
      )
    ),
    xaxis = list(visible = FALSE, range = c(0, 100)),
    yaxis = list(visible = FALSE, range = c(0, 100))
  )
```

```{r, echo=FALSE, warnings=FALSE}
# Render the plot
plot

@@

:::{.border}
```{=html}
<div style="position: relative; display: inline-block;">
  <img src="LipidProbe/Structures/SphingosineStructure.png" alt="Interactive Image" style="width: 500px;">

  <!-- Popup 1 -->
  <div style="position: absolute; top: 20%; left: 10%; cursor: pointer;">
    <span class="popup-hover" title="Popup for Point 1">🔴</span>
  </div>

  <!-- Popup 2 -->
  <div style="position: absolute; top: 70%; left: 50%; cursor: pointer;">
    <span class="popup-hover" title="Popup for Point 2">🔵</span>
  </div>

  <!-- Hover region 1 -->
  <div style="position: absolute; top: 20%; left: 10%; width: 15%; height: 15%; border: 2px solid red; box-sizing: border-box;"
       title="Region 1 Info"></div>

  <!-- Hover region 2 -->
  <div style="position: absolute; top: 50%; left: 50%; width: 20%; height: 20%; border: 2px solid blue; box-sizing: border-box;"
       title="Region 2 Info"></div>
</div>
</div>

<style>
  .popup-hover:hover::after {
    content: attr(title);
    position: absolute;
    top: -30px;
    left: 0;
    background: white;
    color: black;
    padding: 5px;
    border: 1px solid #ccc;
    border-radius: 5px;
    white-space: nowrap;
    box-shadow: 0px 4px 6px rgba(0, 0, 0, 0.1);
  }
</style>
```
:::

@@

### Diazirine: A photo-activated crosslinker

>>>>

---
title: "Multifunctional Lipid Probes Background"
author: "Gaelen Guzman"
format: 
  html:
    embed-resources: false
standalone: true
crossref: 
  ref-hyperlink: true # (default is true)
---

**This repository is still under construction, please forgive us for incomplete sections as we build features!**

@@

* Because at its core biology is just complex chemistry, all biological reactions are mediated through the physical interaction of two or more molecules

* Lipid::Lipid interactomes can provide insight into which of these physical contacts mediate these membrane properties. Lipids are most commonly enmeshed within membrane bilayers and lipid-lipid interactions confer membrane properties such as rigidity or detergent resistance.

* In recent decades, many select lipid species have gained recognition for being crucial signaling molecules. For one example, the sphingolipid sphingosine-1-phosphate (S1P) is potently recognized by an array of S1P receptors which cause T cells to extravasate and chemotax toward sites of infection.

* Understanding the proteins that a given lipid interacts with can reveal great insight into the wide-ranging cellular signaling effects that a lipid can have.

@@

For these reasons, synthetic analogs to native lipids have been a great utility toward probing the biological function of select lipid species. The chief modifications described in this repository (highlighted in the figure below) are: 
1. A diazirine ring 
2. A terminal alkyne
3. A coumarin cage.

These modifications allow for the stabilization of protein::lipid interactions and the affinity enrichment of protein::lipid conjugates, all with tight spatiotemporal control. These modifications are described in greater detail below.

:::{.border}
```{r setup, echo=FALSE, include=FALSE}
library(plotly)
library(tidyr)

@@

# Define hover regions (rectangles) with coordinates based on the top of the image
hover_regions <- data.frame(
  x_start = c(0.0, 0.7),  # Left side of rectangles (fraction of width)
  x_end = c(0.4, 0.9),    # Right side of rectangles (fraction of width)
  y_top = c(0.2, 0.3),    # Top side of rectangles (fraction of height, from top of image)
  y_bottom = c(0.4, 0.4), # Bottom side of rectangles (fraction of height, from top of image)
  text = c("Region 1 Info", "Region 2 Info")
)

# Create grid lines at every 10% for both x and y
grid_lines <- list()
for (i in seq(0, 1, by = 0.1)) {
  # Add vertical grid line
  grid_lines <- c(grid_lines, list(
    list(
      type = "line",
      x0 = i, x1 = i,
      y0 = 0, y1 = 1,
      line = list(color = "gray", width = 1, dash = "dot")
    )
  ))

  # Add horizontal grid line
  grid_lines <- c(grid_lines, list(
    list(
      type = "line",
      x0 = 0, x1 = 1,
      y0 = i, y1 = i,
      line = list(color = "gray", width = 1, dash = "dot")
    )
  ))
}

# Create the plot
plot <- plot_ly()

# Add each rectangle as a separate trace
for (i in 1:nrow(hover_regions)) {
  region <- hover_regions[i, ]
  x_coords <- c(region$x_start, region$x_end, region$x_end, region$x_start, region$x_start)
  y_coords <- c(1 - region$y_top, 1 - region$y_top, 1 - region$y_bottom, 1 - region$y_bottom, 1 - region$y_top)

  plot <- plot %>%
    add_trace(
      x = x_coords,
      y = y_coords,
      type = "scatter",
      mode = "lines",
      line = list(
        color = "blue",     # Border color
        width = 2           # Border width
      ),
      fill = "none",         # No fill color inside the rectangle
      hoverinfo = "text",    # Show text on hover
      text = region$text
    )
}

# Add the image
plot <- plot %>%
  layout(
    images = list(
      list(
        source = image_url,
        xref = "paper",  # Normalized to percentage
        yref = "paper",  # Normalized to percentage
        x = 0,           # Start at 0% of the plot width
        y = 1,           # Start at 100% of the plot height
        sizex = 1,       # Cover 100% of the plot width
        sizey = 1,       # Cover 100% of the plot height
        xanchor = "left",
        yanchor = "top",
        layer = "below"
      )
    ),
    xaxis = list(visible = FALSE, range = c(0, 1)),  # Normalized from 0 to 1
    yaxis = list(visible = FALSE, range = c(0, 1)),
    shapes = grid_lines # Add the grid lines to the layout
  )

```

```{r, echo=FALSE, warnings=FALSE}
#| height: 500px
#| width: 100%
# Render the plot
plot

@@

@@

### Diazirine: A photo-activated crosslinker {#sec-DiazirinePlacement}
